### 2018-02-08 Chapters/Conclusion.tex
Update on Overleaf.

\chapter{Conclusion}
\bigskip

\noindent If anything good can ever be said about the second world war, it might be this: the war effort sparked a massive number of scientific fields.