### 2018-02-08 FrontBackmatter/Abstract.tex
Update on Overleaf.

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}
Short summary of the contents in English\dots

\vfill

\pdfbookmark[1]{Resumo}{Resumo}
\chapter*{Resumo}
Aqui entre o resumo\dots

\endgroup

\vfill