### 2018-02-20 FrontBackmatter/Acknowledgments.tex

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
	Showing gratitude is one of the simplest\\
    yet most powerful things humans can do\\
    for each other.} \\ \medskip
    --- Randy Pausch, The Last Lecture
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

%I must, first and foremost, give thanks for the complete and unwavering support and love from all my family. My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart. Thank you for your patience and for understanding that I was physically far away but very close in heart. A special thanks to my dad, Reynaldo, who always taught me about life and who first awakened my passion for technology and engineering.

%Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love; and for discussing about this work over and over again. Especially, for understanding my absence.

%Thanks to my awesome advisor, Alexandre Linhares, who became a dear friend during this journey.

%Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, André Luiz, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends.

%Thanks to Celene.

%Thanks to Claudio Abreu.

%Thanks to professor Luiz Antônio Joia, who trusted me in the first place. Additionally, thanks to EBAPE / FGV for their excellent teachers and infrastructure.

Finally, thanks to everyone who helped me in any way, for each word, for caring, for your support in difficult moments and for all jokes and talks that made my days worthwhile. Surely this work was done thanks to all of you. After all, nobody does it alone!

\bigskip

\endgroup