### 2018-04-12 Chapters/Chapter01.tex
Hathor: splitting Hathor's architecture and analysis of Bitcoin in separate files.

% !TEX root = ../partial-blockchain.tex

%************************************************
%\chapter{An alternative structure to a blockchain: Confirmations without blocks}
%************************************************

% TODO Include mining-pool proof of hashpower statistics.
% TODO A transaction must confirm the outputs which it is spending (?)

% TODO Lightning network, é uma solução de scaling? Parece que não.

\chapter{Introduction}

\bigskip

\begin{flushright}{\slshape
    {Once I understood the logic behind change addresses...\\
    $ripemd160( SHA256( pubkey ) )$... the inneficiency... \\
    the wasted space; the wasted energy... something clicked.\\
    \medskip

    This is military-grade cryptography and full-blown paranoia.\\
    Whoever did it hears voices whispering `they know everything.' \\
    \medskip

    Other than that, it's a get-rich-quick-scheme \\
    that looks exactly like a get-rich-quick-scheme.}
	\\ \medskip
    --- Alex Linhares, EBAPE/FGV

        \bigskip
        \bigskip
        \bigskip
        \bigskip

    {We are watching History being made.\\
    Or History being repeated.}
	\\ \medskip
    --- David Collum, Cornell University, 2013}
\end{flushright}
\bigskip
\bigskip
\bigskip
\bigskip

The main problem with digital money is how to prevent double spending. As the money is digital, and copies can be made \textit{ad nauseam}, what can prevent counterfeiting? That is, what would prevent users from sending copies of the same money to two (or more) people? That is precisely the problem solved by Bitcoin and its underlying blockchain technology. The current solution behind fiat money is having a single issuer, a central bank --- then trusting the financial institutions.

Bitcoin (BTC) is the first digital currency, also known as digital money, internet money, and cryptocurrency. It is the first currency based on cryptography techniques, distributed and decentralized, and with no central bank. Bitcoin is distributed since its ledger is public and is stored in thousands of computers. It is decentralized because there is no authority (or government) who decides its future --- any decision must be accepted by its community. The security of Bitcoin relies on digital signature technology and network agreement. While digital signature ensures ownership, i.e., the funds may only be spent by their owners, and nobody else; the network agreement both prevents double spending and ensures that all processed transactions have sufficient funds. In short, every transaction must spend only unspent funds, must have enough funds available, and must be signed by its owners, authorizing its execution. When all these requirements are met, the funds are transferred.

According to \citet{barber2012bitter}, despite the 30-year literature on e-cash, most of the proposed schemes requires a central authority which controls the currency issuance and prevents double spending. The no central point of trust and predictable money supply together with a clever solution to the double-spending problem is what separates Bitcoin from the previous e-cash philosophies. % literature

Bitcoin provides interesting incentives to all players, namely the users and the miners. On the one hand, users may have incentives to use Bitcoin because of the following: (i) the fees are small and does not depend on the amount being transfered --- but only in the size (in bytes) of the transaction --- (ii) the transfers will be confirmed in a well-known period; (iii) it is not possible to revert an already confirmed transfer, not even with a judicial order; and (iv) and the currency issuance rate is well-known and preset in Bitcoin's rules, which makes the Bitcoin supply predictable and trustworthy, different from fiat currencies which depends on decisions of their central banks --- i.e., it would be virtually impossible to face a hyper inflation in Bitcoin due to currency issuance. On the other hand, miners have incentive to mine Bitcoin because new Bitcoins are found every ten minutes, and they may also receive the fees of unconfirmed transactions. These incentives have kept the Bitcoin network up and running since 2009 with barely any interruptions (99.99\% uptime).

Since 2009, Bitcoin has been growing and becoming more and more used all around the world. It started as an experiment based on a seminal work by \citet{nakamoto2008bitcoin} and expanded to the most important and successful cryptocurrency with a highly volatile \$192 billion market capitalization, as of this writing \citep{coinmarketcapbtc}. There are hundreds of companies investing in different uses of the technology, from exchanges to debit cards, and billions of dollars being invested in the new markets based on Bitcoin's technology.

Despite Bitcoin's huge success, there are still many challenges to be overcome. We will focus on a specific subset of those challenges, namely the scaling, decentralization, and spam challenges. One important challenge that we will skip is to reduce the size of the ledger (or blockchain), which today is around 125GB and is growing at a rate of 4.5GB per month \citep{blockchaininfosize}.

The network must scale to support hundreds of transactions per second, while its capacity is around only eight transactions per second. Thus, the more Bitcoin becomes popular, the more saturated the network is. Network saturation has many side effects and may affect the players' incentive to keep the network running. The transaction fees have to be increased to compete for faster confirmation. The pool of unconfirmed transactions grows indefinitely, which may cause some transactions to be discarded due to low memory space available, as the previously predictable confirmation time of transactions becomes unpredictable.

Bitcoin seems to have the most decentralized network between the cryptocurrencies, even so, there are few miners and mining pools which together control over 50\% of the network’s computing (hash)power. Thus, they have an oversized influence when it comes to changes in the Bitcoin protocol's behavior, and it is also seen as a problem to be solved. The more decentralized, the more trustworthy Bitcoin is.

Generating new transactions in Bitcoin has a tiny cost because one only has to generate the transaction itself, digitally sign it, and propagate in the Bitcoin network. On the one hand, it means that any device is capable of generating new transactions, but, on the other hand, it makes Bitcoin susceptible to spam attacks. One may generate hundreds of thousands of new valid transactions, overloading the unconfirmed transactions pool and saturating the network.

The number of ideas and publications focusing on improving Bitcoin's design and overcoming those challenges is increasing every day. Many of these proposals are organized into BIPs (Bitcoin Improvement Proposals) which are discussed and implemented by the community; while others come in the form of whitepapers and alternative software forks (which would include the need of a protocol upgrade). Other proposals are published in blogs and forums, describing new cryptocurrencies. Bitcoin's community hardly ever publishes their ideas in academic journals, preferring instead, of BIPs, white papers, and web discussions.

After the launch of Bitcoin, more than 1,000 other cryptocurrencies have been created \citep{coinmarketcap}. In general, they are Bitcoin-like, which means they use similar technologies, including the blockchain. Some cryptocurrencies differs a lot from Bitcoin, like the ones which use the DAG model \citep{dagdiscussion2014, tangle2016, dagcoin2015, sompolinsky2013, lewenberg2015, vorick2015}. We are especially interested in one of them: Iota.

Iota uses a DAG model, called tangle, which has a different design than Bitcoin's blockchain. It has neither mining nor confirmation blocks and transaction fees. Each transaction has its own proof-of-work and is used to confirm other transactions, forming a directed acyclic graph of transactions. Thus, a transaction is said to be confirmed when there are enough proof-of-work from the transactions confirming it directly or indirectly. There is no other way to confirm transactions but generating new transactions.

In Iota, as transactions confirm transactions, the network benefits from a high volume of new transactions. Hence, theoretically, it scales to any large number of transactions per second. The scaling problem of tangle is exactly the opposite of Bitcoin's: it must have at least a given number of transaction per seconds; otherwise, the transactions are not confirmed, and the cryptocurrency does not work. While Iota's network has not reached this minimum number of transactions per second, it uses a central coordinator which works as a trustworthy node \citep{iotacoordinator}. This central coordinator just elucidates that the tangle does not seem to work properly under a low volume of transactions.

Iota's tokens are pre-mined, which means all of them have been issued in the genesis transaction, and there is no token issuance at all. There is no miners and users can only transfer tokens between themselves. This may lead to some incentive issues, because, on the one hand, users would need to keep transferring tokens to keep the network alive, while, on the other hand, users would need to believe that Iota is a good choice to keep using it. As users perception of quality is significantly affected by ``price and congestion'', the \emph{empty-restaurant syndrome} may keep new users away \citep{debo2010prices}.

The present work intends to propose and analyze a new architecture, named Hathor, which lies between Bitcoin and Iota and may be a viable solution to both scaling, centralization, and spam problems. We also present a mathematical analysis of Bitcoin's architecture.

\chapter{Bitcoin \& Blockchain}

In \citeauthor{nakamoto2008bitcoin}'s (2009) seminal paper, there is no distinction between bitcoin and blockchain. They are just one thing which solves an important theoretical problem: how to create a distributed and decentralized digital form of hard money on the internet, in which all users can agree as to whom is entitled to which funds.

But, in practice, it is interesting to separate these concepts. Bitcoin uses the blockchain technology to create a distributed ledger, while the blockchain is a technology which allows information to be stored in an immutable and distributed way.

The blockchain technology works through the creation of new blocks. Each new block confirms that all the previous blocks are valid and have not been tampered with. The mechanism that assures the immutability is proof-of-work, which makes it computationally infeasible to tamper with previous transaction records without having to recalculate all the previous proof-of-works faster than all of the remaining machines of the network. The network agrees that work should be done in the block at the longest chain in the blockchain.

The proof-of-work is a mathematical problem with the following characteristics: (i) it is hard to find a solution; (ii) this hardness level may be adjusted; and (iii) it is fast to check whether the proposed solution is correct.

Bitcoin's blockchain uses the mathematical problem of finding a random number which, after being applied to the hash function SHA-256 twice, results in a number smaller than a given threshold $A$. As SHA-256 is a pseudo-random function, its output is uniformly distributed between 0 and $2^{256}-1$ \citep{gilbert2003security}. Thus, if the given number is $A = 2^{255}$, one has probability 50\% of finding a solution (just the most significant bit of the hash needs to be zero). But if the given number is $A = 2^{240}$, one has probability 0.0015\% of finding a solution (as the ~16 most significant bits of the hash must equal zero). Hence, finding a solution is a hard problem which difficulty depends on the given number $A$. The lesser the given threshold $A$, the higher the difficulty. On the other side, checking whether a solution is correct is fast, one just has to apply the SHA-256 twice and compare.

By design, Bitcoin's blockchain proof-of-work difficulty is dynamically adjusted every 2016 blocks to keep an average pace of 10 minutes between block creation. Thus, the goal is to adjust the difficulty every 14 days. If it takes less than 14 days to find 2016 blocks, that means the network's hash power has increased; thus the difficulty is increased. If it takes more than 14 days to find 2016 blocks, it means the network's hash power has decreased; thus the difficulty is decreased.

When miners are finding a solution to a new block, they are mining or working in the new block. A block is found when a solution to the proof-of-work is found. When a new block is found, it indirectly confirms all the previous blocks in the chain and their transactions. It may happen that two miners find two different blocks in a small interval of time. In this case, both miners will propagate their blocks, and the network will randomly choose one of them as the next block. This phenomenon is called a fork. Thus, when the next block is found, one of those blocks will be confirmed, and the other will be ignored and referred to as an orphan block. As blocks confirm transactions, all the transactions in the orphan block which have not been confirmed by another block will return to the unconfirmed transaction pool. Hence, it is not safe to accept a transaction when it is confirmed by only one block. It is the idea behind the rule of thumb of waiting for at least six confirmations before accepting a transaction.

Newly propagated blocks are validated by the Bitcoin's network. If the solution of the proof-of-work is incorrect or if any transaction included in the block has any issue, then the block is discarded. In order to work properly, the whole network must agree in what is allowed and what is not. Should one think that something should be allowed and accept it in their blocks, the remaining of the network will discard their newly propagated blocks. That is why Bitcoin's network is distributed and decentralized. Everything depends on the agreement of the network, or, precisely, the agreement of the owners of at least 50\% of the hash power. Even if the remaining 49\% disagrees, the 50\% or more who agree will generate, on average, more blocks than the remaining of the network and their rules will prevail on the longest chain. If a disagreement between miners' rules happens, that is referred as either a hard-fork or a soft-fork (in general, a hard-fork relaxes the constraints, while the latter hardens them).

A practical example of a network disagreement is the increase of the block size. No group with more than 50\% of the network's hash power agreed into increasing the maximum block size to increase the number of transactions confirmed by a block and thus increasing the network's capacity. Hence, the capacity remains the same, and the community has been discussing the issue in search of a consensus.

Bitcoin uses the blockchain technology to create a distributed ledger. It allows every new block to generate new bitcoins and also to collect the fees from the confirmed transactions within the block. Bitcoin's transactions have two main parts: (i) inputs, and (ii) outputs. Each transaction sends bitcoins from one or more input addresses to one or more output addresses. In order to prove that one is the owner of the input bitcoins, one must digitally sign the transaction proving such ownership.

The digital signature scheme used by Bitcoin is based on a pair of private and public keys. The private key is used to sign the transaction, while the public key is used to check whether the signature is valid. Thus, the owner of some Bitcoin funds is, in fact, the owner of a pair of private and public keys. The private key must never be publicly published, as whoever has access to the private key is able to spend its funds. In other words, in order to protect their funds, the owners must protect the private key. If one loses their private key, unfortunately, access to their funds will be lost forever. The public key may be used to a proof-of-ownership, i.e., one may publish the public key with some message digitally signed by the private key, proving that he/she is the owner of the funds.

Bitcoins (BTCs) owned by someone are, in fact, unspent outputs in one or more transactions. For instance, one may have 6 BTCs spread between three transactions' unspent output: the first with 1 BTC, the second with 2 BTCs, and the third with 3 BTCs.

In a transaction, the inputs are pointers to other transactions' outputs (which they are spending). A transaction output may only be spent once and thus may not be partially spent. For instance, when one has 3 BTCs in one transaction output and would like to send 1 BTC to a friend, they have to create a transaction with one input spending the 3 BTCs and two outputs, one for the friend with 1 BTC and one's change with 2 BTCs.

Each transaction's output has a script that is executed by the miners to check whether one has or has not permission to spend that output. In other words, whether one has the ownership of that output. In order to execute these scripts, the miners also need some data. This data is given by the transaction which is spending the output.

The output's scripts usually checks whether the public key is valid and whether the digital signature was signed by the private key associated to that public key. Although there are only 3 commonly used scripts, one may create a custom script using the Bitcoin's script language \footnote{Your courageous author once tried to make a transaction with custom script to try to double spend a deposit to an exchange, only to learn through this intrepid adventure that Bitcoin allows only 3 script patterns and the others are treated as invalid.}.

The input contains the data which prove that the sender is the owner of the referred outputs, i.e., the input which must be accepted by the output scripts being spent. Usually, each input has the public key of the sender and a digital signature.

Those users accustomed to block explorers may have been misled by the transaction information that these websites provide.  For example, suppose a miner receives a transaction with ``one input from address A1''.  This ``one input'' actually consists of a pointer to a previous unspent output (i.e., there is no ``input'' address, as is displayed, but only a pointer). This pointer reference allows lookup to be executed in $O(1)$ time.

After lookup, the miner knows how many BTC tokens are available at that unspent output.  But, in order to certify ownership of that output, the miner receives instructions in the form of a script with the rules that lead to the desired unspent output address (and this one is displayed as the input address by those websites). Because this process requires a digital signature, only the holder of the corresponding private key is able to sign such transaction.

Next, we will do a mathematical analysis of Bitcoin in order to better understand its minings properties, how a fork would affect the network and its security against attackers.

\chapter{Analysis of Bitcoin}
\input{./Chapters/hathor-bitcoin-math.tex}

\chapter{Iota \& Tangle}

Iota's underlying technology is Tangle, which has DAG-based architecture with a whole different approach to confirmations. It proposes that there is no need for a block to confirm transactions, as transactions can confirm themselves. Here, each transaction has its own proof-of-work, named weight, and they must confirm two other previous transactions. In this sense, instead of a chain of blocks, the transactions and their confirmations form a directed and acyclic graph (DAG), as in Fig. \ref{fig-tangle-example}.

Like Blockchain, Tangle is another technology to store immutable data and may be the underlying technology to different applications, such as cryptocurrencies, digital contracts (Ethereum-like), digital notaries, and so forth.

Transactions may be either confirmed or unconfirmed. The confirmed transactions have been already confirmed by at least one more transaction. It does not mean they are already irreversible and protected against a double spend attack --- it just means at least one transaction has done some work to confirm it. The unconfirmed transactions are called tips and they are eager to be confirmed. Usually a new transaction selects two tips to confirm, but this rule may not be followed.

Different from Bitcoin, transactions does not have scripts to check whether one may spend the tokens. Instead, it uses a fixed Winternitz hash-based digital signature \citep{dods2005hash}, i.e., whoever correctly signs the transaction may tranfer the tokens. It also supports multi-signature scheme (see \cite{iotamultisign}). Not allowing scripts is a disadvantage when compared to Bitcoin because future applications are limited and changing it would require a major modification in transaction format.

At the beginning, the digital signature algorithm relied on the Curl hash function, which is a ternary hash function designed by Iota's developers. This hash function was replaced by the Kerl hash function after \cite{heilman2017iota} has found a critical vulnerability which enabled practical signature forgery attacks. The Kerl hash function is a variation of SHA-3 also designed by Iota's developers \citep{iotakerl}. Like the Curl hash function, Kerl has not been deeply studied by cryptography researchers and may have critical vulnerabilities.

Iota's tokens are pre-mined, which means they have been issued in the genesis transaction and no more tokens will ever be issued. This means there are no miners in Iota's network and only the users keep the network alive through their new transactions. If, for any reason, the number of transactions per second plummet, the time to confirmed the most recent transactions will increase significantly. This possibly poses a serious risk to the network stability and users' trustworthy.

As there are no miners, there are also no fees, which is a major incentive to new comers. Tokens may be freely transfered without any loses, even for tiny amounts, enabling micropayments. For example, online workers paid per hour may receive their payments every hour, reducing the default risk. This allows untrusted parties to work together reducing the risk for both. The contractor always wants to pay in the end, while the workers always wants to get paid in the beginning.

Not having fees is also also an important feature to the Internet-Of-Things (IoT) technologies. The IoT is a network of smart devices that are connected to the internet and exchange data. It allows, for instance, your refrigerator to do your groceries and automatically pay using your Iota tokens. It also allows your electric car to automatically pay for recharge. For further information, see \cite{fleisch2010internet}.

Theoretically, Iota's network benefits from high volume of transactions. The more transactions are coming, the faster previous transactions are confirmed. This is a major feature of Tangle which contrasts with Bitcoin's difficulty to scale. On the other hand, low volume of transactions is a primary problem for Iota, since it would take too long to confirm previous transactions. As it is a known problem, Iota has created a central coordinator which works as a trustworthy node, clearing transactions. In other words, every transaction directly or indirectly confirmed by the central coordinator is assumed to be cleared. Iota's developers claims that the coordinator will be turned off when the network outgrows a minimum (unknown) size.

An important factor of Iota is how the new transactions choose which transactions they will confirm. There are several possible approaches, such as randomly selecting two of the unconfirmed transactions (tips). In Fig \ref{fig-tangle-example}, the reader may have noticed that transaction 8 will confirm transaction 4, which has already been confirmed by transaction 7. It may be on purpose, or maybe transaction 4 was unconfirmed when it was chosen, but it got confirmed during the calculation of the proof-of-work or the network propagation of the transaction. The selection algorithm seems to be important to protect the network against double spend attacks. For further information, see \cite{tangle2016}.

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-example.pdf}
\caption{White nodes represent transactions that have been confirmed at least once. Green circles represent unconfirmed transactions (tips). Gray and dashed nodes are the transactions currently solving the proof-of-work in order to be propagated.\label{fig-tangle-example}}
\end{figure}

Transactions have an accumulated weight which may be interpreted as how hard it is to rollback a transaction. It is analogous to the number of confirmations of a block in Bitcoin. The higher the accumulated weight, the safer the transaction. Let $A$ be a transaction, its accumulated weight is the sum of all weights of the transactions which confirm $A$, including A itself, i.e., $w_A + \sum_{A \leadsto P} w_P$. For example, in Figure \ref{fig-tangle-example}, the accumulated weight of transaction 3 is the sum of the weights of the transactions 3, 5, 6, 7, and 8.

The score of a transaction is a measure of how much proof-of-work has been done before the transaction has been created. As the heighest score of the network increases over time, comparing a transaction's score with the highest score of the network indicates the ``age'' of the transaction. The score of a transaction A is the sum of all weights of the transactions which are being confirmed by A, including A itself, i.e., $w_A + \sum_{P \leadsto A} w_P$. For example, in Fig. \ref{fig-tangle-example}, the score of transaction 3 is the sum of the weights of the transactions 1, 2, and 3.

Another measure of the ``age'' of a transaction is its height. The height of a transaction A is the length of the longest path from transaction A to the genesis transaction. For example, in Fig. \ref{fig-tangle-example}, the height of transaction 5 is four ($5 \rightarrow 3 \rightarrow 2 \rightarrow 1 \rightarrow$ genesis). The lower the height, the older the transaction.

The depth of a transaction A is a measure of the youth of the transaction. It is the length of the longest path in the inverted graph from transaction A to any unconfirmed transaction (tip). For example, in Fig. \ref{fig-tangle-example}, the depth of transaction 2 is three ($2 \rightarrow 3 \rightarrow 5 \rightarrow 8$). It is the opposite of the height. The lower the depth, the younger the transaction. When a new transaction is confirming two transactions with high depth, it is referred to as lazy transaction.

The higher the volume of new transactions, the more unconfirmed transactions will appear. In Fig. \ref{fig-tangle-swarm}, the reader can notice that the number of new transactions was increased for a while, and then decreased back to the original value. The Iota has behaved well when exposed to a high load scenario, since it reduced the number of tips to only three after the high demand has ceased. It is like a moving swarm which gets wider when the number of new transactions increases and gets thinner when the number of new transactions decreases.

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-swarm.pdf}
\caption{Suddenly the number of transactions per second increases and the width of the swarm grows. After a while, the number of transactions per second decreases and the width of the swarm shrinks.\label{fig-tangle-swarm}}
\end{figure}

Conflicting transactions may happen when two or more transactions try to spend the same tokens --- or, in the Bitcoin's transaction format, try to spend the same output. In this case, the network must choose which of the transactions will be accepted and the other one will be invalidated, even when both have already been confirmed. In fact, when one transaction is invalidated, the whole sub-DAG which confirms it is also invalidated. In this case, it may happen to reverse some transactions.

Intuitively, when there is a conflict, the network should accept the transaction which has greater accumulated weight, invalidating the others (see Fig. \ref{fig-tangle-conflict}). But it may be not enough to prevent some attacks like the nuclear submarine attack.

\begin{figure}[ht]
\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-conflict.pdf}
\caption{The red nodes are transactions which had some conflict with previous transaction and were invalidated by the network. Notice that none of them have been confirmed.\label{fig-tangle-conflict}}
\end{figure}

The nuclear submarine attack, also known as the parasite chain attack, is when the attacker generates a separate DAG (or a side DAG), with many transactions and a lot of proof-of-work. This side DAG is off the network, i.e., its transactions have not been propagated. Then, at a convenient moment, the attacker suddenly propagates these transactions. The whole network needs to decide how to handle these transactions.

If the transactions have no conflict with any transaction of the main DAG, i.e., there is no transaction spending the same tokens, then it is easy to handle the transactions. But, as it is an attack, there will be some conflicts, and it is not easy to choose which transaction should be invalidated. As the attacker has been generating a separate DAG, the conflicting transaction may have an accumulated weight similar or greater than the transaction which is already in the main DAG. Hence, using only the accumulated weight may not be enough to prevent this attack.

For example, the attackers generate (and do not propagate) a transaction that transfers all their funds to another address. Then, they start to generate many new transactions which confirm themselves and even confirm some of the transactions in the main DAG, but none of these transactions are also propagated to the network. Afterward, the attacker buys something in the real world, pays with cryptocurrency, and wait until the payment gets the accumulated weight demanded by the merchant. Finally, the attacker suddenly propagates all the transactions to the network in a small window of time. If the criteria is to validate the transaction with higher accumulated weight, the network will accept the attackers' original transaction instead of the one used to pay the merchant. Hence, the merchant transaction is invalidated, and the double spend attack has succeeded.

By default, Iota uses the Markov Chain Monte Carlo (MCMC) algorithm to select the two tips. For further information about attacks and strategies to prevent them, including the MCMC algorithm, see \cite{tangle2016}.

\chapter{Hathor's architecture}
\input{./Chapters/hathor-architecture.tex}

\chapter{Methodology}

The methodology we have used is computer simulation. Through the simulation of many scenarios of Hathor, we will understand how the network behaves in complex scenarios, including when the load suddenly increases, and when the network is under attack.

The simulator has been developed using an event-based design which is capable of running hours of simulation in just a few minutes. It creates agents who decide to make a transaction, then they select which transactions will be confirmed, then they spend some time working in the proof-of-work, and, finally, they propagate the transaction to the network. The other agents receive the transaction and may accept or deny it. The agents may use different parameters among themselves.

When a new transaction emerges, it chooses two tips to confirm before solving the proof-of-work. When it finishes solving the proof-of-work, it is propagated and becomes a tip. So, two new transactions may choose the same tips to confirm. If there are $t$ tips, a new transaction will randomly choose 2 out of these $t$ tips, even if they have already been chosen by other new transactions --- in fact, they do not know which have been chosen because these new transactions have not being propagated yet.

When a new transaction is added to the Hathor's network, it uses a depth-first search \citep{cormen2009introduction} (i) to update the aggregated weight of the directly and indirectly confirmed transactions, (ii) to calculate its own score, and (iii) to generate a topological sort \citep{cormen2009introduction}. The topological sort is used to calculate the longest path from all other transactions to the new transaction, and this longest path is used to update the depth of the whole DAG. If the new transaction confirms only transactions which has already been confirmed, the depth update is skipped thanks to Theorem \ref{theorem-new-tx-not-tip}.

The amount of time spent to solve the proof-of-work is a sample from an exponential distribution.

The time between two new transactions is sampled from an exponential distribution with parameter $\lambda$. The value of $\lambda$ is changed over time, causing changes in the load of the network. We consider 1 tx/s to be a low load, and 15 tx/s to be a high load. Even though we have already tested up to 100 tx/s.

The simulator will output different reports: (i) snapshots of the DAG at interesting moments, such as Fig. \ref{fig-tangle-example}, \ref{fig-tangle-swarm}, and \ref{fig-tangle-conflict}; (ii) histogram of confirmation time, such as Fig. \ref{fig-tangle-hist} and \ref{fig-tangle-hist-2}; and many others.

\chapter{Analysis of Hathor}

Hathor's architecture lies between Iota and Bitcoin's architectures. It is similar to Bitcoin's architecture when the number of transactions per second is low, while it is similar to Iota's architecture when the number of transactions per second is high. In order to check this statement, I have run some simulations.

First, two simulations with the extreme cases: (i) no miners, (ii) no transactions. The first should have exactly the same behavior as Bitcoin's blocks with one block after the other forming a long chain. The latter should have a similar behavior as Iota's structure, forming a Direct Acyclic Graph (DAG) with transactions confirming transactions. We can notice in Figure \ref{fig:hathor-similarities} that both cases seem to be correct.

Then, I have run a simulation with both miners and transactions. As we can see in Figure \ref{fig:hathor-dag}, Hathor's structure is a mix of Iota and Bitcoin's structure. The transactions are forming a DAG while, in parallel, the blocks are forming a chain. Differently from Bitcoin, there is no orphan block which means some blocks may be found close to the previous one.

\begin{figure}[!htb]
\centering
\subfloat[No miners]{\includegraphics[width=0.5\textwidth]{./images01/sim/no_miners.pdf}}
\subfloat[No transactions]{\includegraphics[width=0.5\textwidth]{./images01/sim/only_miners.pdf}}
\caption{Visualization of a Hathor's graph in two particular cases: (a) no miners, (b) no transactions. It shows that when there are no miners, Hathor is similar to Iota (same structure, but different parameters), and when there is no transactions, it is similar to Bitcoin.\label{fig:hathor-similarities}}
\end{figure}

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{./images01/sim/hathor.pdf}
\caption{Visualization of a Hathor's graph with transactions and blocks. Red boxes are blocks, and white circles are simple transactions. The arrows show the confirmations.\label{fig:hathor-dag}}
\end{figure}

Another evidence that Hathor lies between Iota and Bitcoin is found when comparing the time to confirm a transaction. In this context, a transaction is said to be confirmed when it has reached an accumulated weight similar to six times the hash rate of the whole network (miners and new transactions). This criteria is equivalent to the well-known ``6 confirmations'' of Bitcoin, which is adopted by almost the whole network.

Thus, I have run a simulation in which miners are majority and there are few transactions. In Figure \ref{fig:hathor-tct-low-mid}, we may see a good fit between the confirmation time of a transaction and the theoretical distribution of the time to find six blocks in Bitcoin (which is $Y_6$ and follows an Erlang distribution). When the load is increased, Hathor's confirmation time is reduced and diverges from Bitcoin's distribution. The reasoning is that confirmations coming from other transactions start to play an important role and accelerates the speed of confirmations.

\begin{figure}[!htb]
\centering
\subfloat[Low load]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-low-load.png}}
\subfloat[Mid load \label{fig:hathor-tct-mid-load}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-mid-load.png}}

\caption{Confirmation time in two scenarios: (a) low load, (b) mid load. The red curve is the distribuion of the time to find six blocks in Bitcoin (which follows an Erlang distribution). As we can notice, in the low load scenario, Hathor's confirmation time behaves just like Bitcoin's. When the load is increased, it starts to diverge from Bitcoin's distribution. \label{fig:hathor-tct-low-mid}}
\end{figure}

We may see Hathor's confirmation time moving from Bitcoin's to Iota's in Figure \ref{fig:hathor-tct-many}. Notice that the confirmation timer is getting smaller as the number of transactions per second increases. Figure \ref{fig:hathor-tct-many-right} is a zoom-in in the right side, and we can see again the good fit between Hathor Bitcoin's confirmation time under low load.

\begin{figure}[!htb]
\centering
\subfloat[The whole picture]{\includegraphics[width=\textwidth]{./images01/sim/tct-many-loads-2.png}}

\subfloat[Zoom in the left part of the above chart]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-many-loads-3.png}}
\subfloat[Zoom in the right part of the above chart \label{fig:hathor-tct-many-right}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-many-loads-4.png}}

\caption{Confirmation time in many scenarios, moving from a low load ($\lambda_\text{TX} = 0.015625$) to a high load ($\lambda_\text{TX} = 2$). \label{fig:hathor-tct-many}}
\end{figure}

But, what would happen if, instead of changing the number of transactions per second, we change the relative hash power between miners and transactions? In previous simulations, the miners had a hash rate in the same magnitude as the transactions. In Figure \ref{fig:hathor-tct-100k-10k}, we can see the same simulation as in Figure \ref{fig:hathor-tct-mid-load}, but with the miners' hash rate ten times the transactions'. Besides the difference in the shape of the distribution, we can see that it is moving back towards Bitcoin's confirmation time distribution. It also makes sense because increasing the miners' hash rate increases the required minimum accumulated weighted for confirmed transactions. Therefore, more transactions are necessary to give more accumulated weight. As the number of transactions per second was not changed, most of the work of confirmations were done by blocks (and not by transactions). To confirm this idea, I kept the miners' hash rate ten times the transactions' and increased 16 times the number of transactions per second. As we can see in Figure \ref{fig:hathor-tct-100k-10k-high}, when the number of transactions per second is increased, its role in the accumulated weight also increases and it goes farther from Bitcoin's distribution.

\begin{figure}[!htb]
\centering
\subfloat[Same load as in Figure \ref{fig:hathor-tct-mid-load}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-100k-10k.png}}
\subfloat[High load, 16 times of (a)\label{fig:hathor-tct-100k-10k-high}]{\includegraphics[width=0.5\textwidth]{./images01/sim/tct-100k-10k-high-load.png}}

\caption{Confirmation time with miners' hash rate ten times the transactions'. \label{fig:hathor-tct-100k-10k}}
\end{figure}

Finally, even with both blocks and transactions, Hathor's blocks are similar to Bitcoin's blocks, and they share the same math. To confirm that, see Figure {fig:hathor-tbb}, where the red curve is Bitcoin's distribution of time between blocks and the blue histogram is Hathor's time between blocks. I also made several tests adding and removing miners to check the difficulty adjustment and it worked properly.

\begin{figure}[!htb]
\centering
\subfloat[1 miner]{\includegraphics[width=0.5\textwidth]{./images01/sim/tbb-1.png}}
\subfloat[2 miners after difficulty adjustment]{\includegraphics[width=0.5\textwidth]{./images01/sim/tbb-2-after.png}}

\caption{Histogram of time between blocks. The red curve the Bitcoin's theoretical distribution of time between blocks. As we can notice, the fit is very good. \label{fig:hathor-tbb}}
\end{figure}

Afterward, to a better visualization of a Hathor's network, I run a simulation classifying the transactions in either confirmed, in-progress, or unconfirmed (tips). I also showed the transactions solving the proof-of-work which had not been propagated yet. See Figure \ref{fig:hathor-dag-big} and notice the chain of blocks inside the DAG.

\begin{figure}[!htb]
\centering\includegraphics[width=\textwidth]{./images01/sim/hathor-2.pdf}
\caption{Visualization of a Hathor's graph with transactions and blocks. Red boxes are blocks; green circles are confirmed transactions; white circles are in-progress transactions; yellow circles are unconfirmed transactions (tips); and grey circles are transactions solving the proof-of-work which have not been propagated yet. The arrows show the confirmation chain. Block's arrows are in bold. \label{fig:hathor-dag-big}}
\end{figure}

%The number of transactions solving the proof-of-work may be modeled by a Birth and Death process, since if we have $k$ transactions solving the proof-of-work only two things may happen: either a new transaction will be created or one the transactions will finish solving the proof-of-work. Let the time between new transactions follows an exponential distribution with $\mu$ parameter, then the probability of a new transaction emerges before any transaction finishes its proof-of-work is $q_k = \mathbf{P}(T_{\text{new tx}} = \min\{T_{\text{new tx}}, T_1, T_2, \dots, T_k\}) = \frac{\mu}{\mu + \sum_{i=1}^k \lambda_i}$. Hence, the probability of any transaction finishes its proof-of-work before a new transaction emerges is $p_k = 1 - q_k = \frac{\sum_{i=1}^k \lambda_i}{\mu + \sum_{i=1}^k \lambda_i}$.

\begin{theorem}
\label{theorem-new-tx-not-tip}
When a new transaction confirms an already confirmed transaction, the depth of all transactions remains the same, i.e., the depth of the transactions are only changed when a new transaction confirm an unconfirmed transaction.
\end{theorem}

\begin{theorem}
The height of a transaction is equal to the maximum height of its parents plus one.
\end{theorem}

\begin{theorem}
If new transactions choose randomly the unconfirmed transactions to be confirmed, then no unconfirmed transaction will be left behind, i.e., all transactions will be confirmed in due time.
\end{theorem}

\chapter{Conclusion}

Bitcoin's underlying technology blockchain has been called by many as a major invention, even comparable to the invention of the internet. But it is unlikely that Bitcoin and blockchain have achieved the final or most optimal design for a secure and scalable electronic transaction system. In this work, I proposed and analysed a new architecture named Hathor, which seems a scalable alternative to Bitcoin.

Today, Bitcoin network can barely handle 8 transactions per second without increasing the unconfirmed transaction list to hundreds of thousands --- several transactions take days to be confirmed. In order to increase Bitcoin's capacity, its community has first proposed and implemented segregated witness, which was not enough. Finally, they proposed the lightning network, which is in development and should be available in the next months.

- What is better: solves a PoW with $w$, or solves $n$ PoWs with $w/n$?

%\chapter{Proposal: Questions to be explored}

%Given these preliminaries, these are some questions with which we will be concerned.

%In this paper, we will analyze the network scaling and security. How a Hathor network scales, simulating different loads and measuring the bandwidth, computational effort, and storage space necessary to handle all the transactions in time. What is the minimum transaction's aggregated weight which it may be considered unlikely to be reversed?

%We are interested in how the rate of new transactions affects how long it takes to a new transaction to be confirmed for the first time. We had already run some simulations under normal load (Fig. \ref{fig-tangle-hist}) and high load (Fig. \ref{fig-tangle-hist-2}).

%\begin{figure}[ht]
%\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist.png}
%\caption{Histogram of how long has been a transaction waiting until its first confirmation. It was a simulation of 15 minutes with new transactions rate changing between 1 and 15 tx/s.\label{fig-tangle-hist}}
%\end{figure}

%\begin{figure}[ht]
%\centering\includegraphics[width=\textwidth]{./images01/fig-tangle-hist-2.png}
%\caption{Histogram of how long has been a transaction waiting until its first confirmation. It was a simulation of 5 minutes with new transactions rate of 50 tx/s, i.e., very high load.\label{fig-tangle-hist-2}}
%\end{figure}

%Besides the time to the first confirmation, it is also important to measure how long it takes to a new transaction reach a specific accumulated weight. It is useful for exchanges and merchants to set their minimum requirements. Bitcoin's exchanges and merchants usually requires a minimum of 6 confirmations blocks.

%OLD \cite{tangle2016} suggests that constraining transactions' weight in a range would both prevent spam, because it would be necessary a minimum work to propagate a new transaction, and attacks, because an attacker would not be allowed to propagate a transaction with very high weight. We agree with the spam argument, and would like how a minimum weight would affect mobile devices's new transactions. We partially agree with the upper bound, because an attacker would be able to generate many transaction with lower weight. We will analyze if constraining transactions' weight would be effective against nuclear submarine attacks.

%OLD We have another suggestion to prevent nuclear submarine attacks: set a maximum depth for the transactions confirmed by a new transaction. So, new transactions would have to confirming newer transactions instead of old transactions, and a nuclear submarine attack would be controlled because the attacker would be not be able to create a large separate DAG. But the maximum depth rule would possibly also affect transaction created by low (hash)power devices, because it would take a longer time to solve the proof-of-work and, if the confirmations are going fast, the transaction would possibly be invalidated. This raises another important question: what would be an optimal maximum depth allowed?

%OLD If we would like to include fee transactions, how would it be distributed between confirmations? As transactions may have multiple direct confirmations, it is not obvious who would receive the transaction fee. An intuitive suggestion would be to give the fee to the first confirmation with a minimum weight, but it might have some propagation problems, because it takes a while to propagate a transaction in the network and there would be conflict.

%OLD In order to solve the fee distribution problem, we will simulate how the network would behave if we include blocks. These blocks would be like Bitcoin's, with an adjustable proof-of-work according to the network's hashpower. They would receive the fees from all confirmed transactions which had not been confirmed by any block before, and they would be able to generate new coins. These blocks would also help solving other problems.

%\begin{enumerate}
%\item Which parameters turns the network more or less vulnerable to attacks?
%\item Is there an optimal strategy to attack?
%\item Could new coins be generated demanding a proof-of-work with dynamic difficulty, like in Bitcoin?
%\item What is the equation of number of tips over time? What is the equation of how long it takes to a transaction to be confirmed for the first time? What is the equation of the accumulated weight of a transaction over time? Some equations have already been proposed by \citet{tangle2016}.
%\end{enumerate}