### 2018-04-25 FrontBackmatter/Abstract.tex
Abstract and acnowledgments (incomplete).

%*******************************************************
% Abstract
%*******************************************************
%\renewcommand{\abstractname}{Abstract}
\pdfbookmark[1]{Abstract}{Abstract}
\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax

\chapter*{Abstract}

This thesis presents three specific, self-contained, scientific papers in the Computational Management Science area.
Modern management and high technology interact in multiple, profound, ways. Professor Andrew Ng tells students at Stanford’s Graduate School of Business that ``AI is the new electricity'', as his hyperbolic way to emphasize the potential transformational power of the technology.

The first paper is inspired by the possibility that there will be some form of purely digital money and studies distributed ledgers, proposing and analyzing Hathor, an alternative architecture towards a scalable cryptocurrency.

The second paper may be a crucial item in understanding human decision making, perhaps, bringing us a formal model of recognition-primed decision. Lying at the intersection of cognitive psychology, computer science, neuroscience, and artificial intelligence, it presents an open-source, cross-platform, and highly parallel framework of the Sparse Distributed Memory and analyzes the dynamics of the memory with some applications.

Last but not least, the third paper lies at the intersection of marketing, diffusion of technological innovation, and modeling, extending the famous Bass model to account for users who, after adopting the innovation for a while, decide to reject it later on.

\vfill

\pdfbookmark[1]{Resumo}{Resumo}
\chapter*{Resumo}
Aqui entre o resumo\dots

\endgroup

\vfill