### 2018-04-25 FrontBackmatter/Acknowledgments.tex

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
	Showing gratitude is one of the simplest\\
    yet most powerful things humans can do\\
    for each other.} \\ \medskip
    --- Randy Pausch, The Last Lecture
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

I must, first and foremost, give thanks for the complete and unwavering support and love from all my family. My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart. Thank you for your patience and for understanding that I was physically far away but very close in heart. A special thanks to my dad, Reynaldo, who always inspired me and who first awakened my passion for technology and engineering.

Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love. Especially, for understanding my absence.

I'm very grateful for the friendship, guidance, and invaluable help of my advisor Alexandre Linhares. Without him this work would never be possible. His passion for cognitive sciences, artificial intelligence, and economics has influenced everyone around him. His ideas have profoundly influenced my thoughts and my life since we have met. I must not forget his wife Paula, who became a dear friend and has always supported me as well, even when I was working in her living room until the sunrise. Thank you for this amazing journey, my friends.

Thanks to my dear friends Daniel Chada, Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends. I will never forget our all-you-can-eat adventures in Bros restaurant, the snacks in Catarina, and the drinks in Bigode bar. You have become close friends who I will carry for life.

Thanks to my long-time partner and dear friend Claudio Abreu, who always encouraged me to take risks and to improve myself. He is the most fearless person I have ever met.

Thanks to EBAPE / FGV for their great faculty, staff, and infrastructure. A special thanks to Celene Silva Melo for her empathy and support besides all the help with paperwork, funding, and many other things. Another special thanks to doctor Luiz Antônio Joia, who trusted me in the first place and guided my first steps into the world of academy.

Finally, thanks to everyone who helped me in any way, for each word, for caring, for your support in difficult moments and for all jokes and talks that made my days worthwhile. Surely this work was done thanks to all of you. After all, nobody does it alone!

\bigskip

\endgroup