### 2018-05-10 defense-presentation.tex
Add defense presentation.

\documentclass{beamer}

\usetheme{Boadilla}

\title{Essays in Computational Management Science}
\author{Marcelo Salhab Brogliato}
\institute{EBAPE / FGV}
\date{\today}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}
\frametitle{Outline}
\begin{enumerate}[I]
\item Hathor: An alternative towards a scalable cryptocurrency
\item An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics
\item Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users
\end{enumerate}
\end{frame}

\part{Hathor: An alternative towards a scalable cryptocurrency}
\begin{frame}
\frametitle{Hathor: An alternative towards a scalable cryptocurrency}

The main contributions are:
\begin{enumerate}[i]
\item Mathematical analysis of Bitcoin
\item Proposal of Hathor
\item Analysis of Hathor
\end{enumerate}
\end{frame}

\part{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}
\begin{frame}
\frametitle{An invitation to Sparse Distributed Memory: from the theoretical model to the system dynamics}

The main contributions are:
\begin{enumerate}[i]
\item A validated, open-source, cross-platform framework
\item Identification and modeling of an autocorrelation
\item Noise filtering application
\item Classification application
\item Reinforcement learning application
\end{enumerate}
\end{frame}

\part{Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users}
\begin{frame}
\frametitle{Diffusion and dismissal of innovation: forecasting the number of Facebook’s active users}

The main contributions are:
\begin{enumerate}[i]
\item Add a rejection parameter to the model
\item Estimate the parameters using only the number of active users
\item The possibility of reducing the number of active users (instead of only increase)
\end{enumerate}
\end{frame}

\end{document}