### 2018-05-11 FrontBackmatter/Acknowledgments.tex
Changes in Hathor's bitcoin math, acknowledgments, and bass model.

%*******************************************************
% Acknowledgments
%*******************************************************
\pdfbookmark[1]{Acknowledgments}{acknowledgments}

\begin{flushright}{\slshape
	Showing gratitude is one of the simplest\\
    yet most powerful things humans can do\\
    for each other.} \\ \medskip
    --- Randy Pausch, The Last Lecture
\end{flushright}

\bigskip

\begingroup
\let\clearpage\relax
\let\cleardoublepage\relax
\let\cleardoublepage\relax
\chapter*{Acknowledgments}

I must, first and foremost, give thanks for the complete and unwavering support and love from all my family. My beloved wife Patrícia, my wonderful parents Reynaldo and Angelina, my lovely sister Flávia, my sweet grandmas Geny and Rosaria, and all others who will always be remembered in my heart. Thank you for your patience and for understanding that I was physically far away but very close in heart. A special thanks to my dad, Reynaldo, who always inspired me and who first awakened my passion for technology and engineering.

Thanks to my wife, Patrícia, for sharing the dreams, the passion, and the craziness, for all her caring and love. Especially, for understanding my absence and my unusual working times. Her loyalty and commitment make me the luckiest man alive. I love you.

I'm very grateful for the friendship, guidance, and invaluable help of my advisor Alexandre Linhares. Without him this work would never be possible. His passion for cognitive sciences, artificial intelligence, and economics has influenced everyone around him. His ideas have profoundly influenced my thoughts and my life since we have met. His optimism and trust in my abilities were undying. I must not forget his wife Paula, who became a dear friend and has always supported me as well, even when I was working in her living room until sunrise. Thank you for this amazing journey, my friends.

Thanks to my friends Layla Mendes, Renato Kogeyama, André Luiz, Luiz Sacramento, Kaillen Givigi, Jamil Chivitarese, Lucia Barros, Daniel Modenesi, Bernardo Machado, and Andréia Sodré, always available for serious talking, laughing, and sharing their wide and deep knowledge, even at lunchtime and weekends. I will never forget our all-you-can-eat adventures in Bros restaurant, the snacks in Catarina, and the drinks in Bigode's bar. You have become close friends who I will carry for life.

Thanks to my long-time business partner and friend Claudio Abreu, who share a deep delight for engineering and has taught me so much about life, family, and business. He has always put faith in me and cheered for my victories. Thank you, my dear friend.

I thank my dear buddy and coauthor Daniel Chada for sharing the passion for research, always inspiring and amusing me with his astonishing ideas. Our long discussions have definitely shaped my researching abilities. Thank you, bro!

Thanks to EBAPE / FGV for their finest faculty, staff, and infrastructure. A special thanks to Celene Silva Melo for her empathy and support besides all the help with paperwork, funding, and many other things. Another special thanks to professor Luiz Antônio Joia, who trusted me in the first place and guided my first steps into the world of academy.

Last, but far from least, I thank Kallen Givigi for being the ``Psych-os'' lab's protector and defender. She has always been there solving all sort of things (even personal matters) and whipping our lab into shape. Always kind and smiling, she gently asked for more focus and less mess. Your request is my command, my lady.

Finally, thanks to everyone who helped me in any way, for each word, for caring, for your support in difficult moments and for all jokes and talks that made my days worthwhile. Surely this work was done thanks to all of you. After all, nobody does it alone!

\bigskip

\endgroup