### 2018-08-22 paperIEEE.tex
wtf no graphics?

%% LyX 2.3.0 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.

\RequirePackage{fix-cm} % fix some latex issues see: http://tug.org/texmf-dist/doc/latex/base/fixltx2e.pdf

\documentclass[twocolumn,journal]{IEEEtran}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage{calc}
\usepackage[unicode=true,
 bookmarks=true,bookmarksnumbered=true,bookmarksopen=true,bookmarksopenlevel=1,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
 {hyperref}
\hypersetup{pdftitle={Your Title},
 pdfauthor={Your Name},
 pdfpagelayout=OneColumn, pdfnewwindow=true, pdfstartview=XYZ, plainpages=false}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
% protect \markboth against an old bug reintroduced in babel >= 3.8g
\let\oldforeign@language\foreign@language
\DeclareRobustCommand{\foreign@language}[1]{%
  \lowercase{\oldforeign@language{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
% for subfigures/subtables
\usepackage[caption=false,font=footnotesize]{subfig}

\usepackage{amsmath,amssymb}
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem*{cor}{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\DeclareMathOperator\erfc{erfc}
\DeclareMathOperator\erf{erf}

\makeatother

\begin{document}

\title{Autocorrelation in Sparse Distributed Memory}

\author{Marcelo S. Brogliato,~\IEEEmembership{Member,~IEEE,} and Alexandre
Linhares, \IEEEmembership{Member,~IEEE,}\thanks{Dr. Brogliato is with sdm.ai and with Behavioral and Decision Sciences,
EBAPE, Fundação Getulio Vargas, Rio de Janeiro, Brazil, e-mail: \protect\href{http://msbrogli@sdm.ai}{msbrogli@sdm.ai}.}\thanks{Dr. Linhares is with sdm.ai and with Behavioral and Decision Sciences,
EBAPE, Fundação Getulio Vargas, Rio de Janeiro, Brazil, e-mail: \protect\href{http://linhares@sdm.ai}{linhares@sdm.ai}.}}

\markboth{Working Paper}{Brogliato and Linhares: Autocorrelation of SDM}

\IEEEpubid{0000\textendash 0000/00\$00.00~\copyright~2012 IEEE}
\maketitle
\begin{abstract}
Sparse Distributed Memory (SDM) is a neuroscientific and psychologically
plausible model of human memory. In this paper we present an anomaly
between its previously theoretized behavior and its actual behavior,
and demonstrate that this anomaly is due to the autocorrelation involved
in the model. Our findings have systemwide implications for those
willing to implement and apply SDM in computational intelligence settings.
\end{abstract}

\begin{IEEEkeywords}
Sparse Distributed Memory, Autocorrelation, Neural Networks, Memory,
Systemwide properties.
\end{IEEEkeywords}

\IEEEpeerreviewmaketitle{}

\section{Introduction}

\IEEEPARstart{S}{parse distributed memory} has been shown to be
fodastic.

In this paragraph we explain what SDM is.

In this other paragraph we explain what the equator distance is. Then,
this introductory section breaks and moves on to the main findings
of this paper.

\begin{figure}[!htb]
\centering\includegraphics{./images02/calculated-table-72.png}
\caption{Kanerva's original Figure 7.3 generated using the equations from \citet{brogliato2014sparse}.
\label{fig:kanerva-figure-73-calculated}
}
\end{figure}

\input{./Chapters/sdm-autocorrelation.tex}

\section{Conclusions}

bla bla

\appendices{}

\section*{Acknowlegment}

The authors would like to thank Pentti Kanerva, Eric Paul Nichols,
José Ricardo de Almeida Torreão, Moacyr Alvim Horta Barbosa da Silva,
Flavio Codeço Coelho, Horacio Hideki Yanasse, and Paulo Murilo Castro
de Oliveira for their careful, comprehensive, reviews of the first
author\textquoteright s theses.

\bibliographystyle{plain}
\bibliography{mybib02}

\begin{IEEEbiography}[{\fbox{\begin{minipage}[t][1.25in]{1in}%
Replace this box by an image with a width of 1\,in and a height of
1.25\,in!%
\end{minipage}}}]{Your Name}
 All about you and the what your interests are.
\end{IEEEbiography}

\begin{IEEEbiographynophoto}{Coauthor}
Same again for the co-author, but without photo
\end{IEEEbiographynophoto}

\end{document}